### 2016-09-27 revision/abstract.tex
Improvements

We design algorithms for online linear optimization that have optimal regret
and at the same time do not need to know any upper or lower bounds on the norm
of the loss vectors. We design them as instantiation of the Follow the Regularized Leader (FTRL) and Mirror Descent (MD) meta-algorithms.
We achieve adaptiveness to the norms of the loss vectors
by scale invariance, i.e., our algorithms make exactly the same decisions if
the sequence of loss vectors is multiplied by any positive constant.  The algorithm based on FTRL works for any decision set, bounded or unbounded.  For unbounded decisions sets, this is the first adaptive algorithm for online linear
optimization with a non-vacuous regret bound.
We also show lower bounds on scale-free algorithms based on MD on unbounded domains.

>>>>

We design and analyze algorithms for online linear optimization that have
optimal regret and at the same time do not need to know any upper or lower
bounds on the norm of the loss vectors. Our algorithms are instances of the
Follow the Regularized Leader (FTRL) and Mirror Descent (MD) meta-algorithms. We
achieve adaptiveness to the norms of the loss vectors by scale invariance, i.e.,
our algorithms make exactly the same decisions if the sequence of loss vectors
is multiplied by any positive constant.  The algorithm based on FTRL works for
any decision set, bounded or unbounded.  For unbounded decisions sets, this is
the first adaptive algorithm for online linear optimization with a non-vacuous
regret bound. In contrast, we show lower bounds on scale-free algorithms based
on MD on unbounded domains.

### 2016-09-27 revision/ada-ftrl.tex
Improvements

Here we show that it is possible to derive scale-free algorithms in an alternative way. We generalize the \textsc{AdaHedge}
algorithm~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to the OLO setting,
showing that it retains its scale-free property. The analysis is very general
and based on general properties of strongly convex functions, rather than
specific properties of the entropic regularizer like in \textsc{AdaHedge}.

@@

Our main result is an $O(\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$ upper bound on
the regret of the algorithm after $T$ rounds, without the need to know beforehand an upper bound on $\|\ell_t\|_*$.  We prove the theorem in
Section~\ref{section:ada-ftrl-regret-bound}.

@@

\begin{lemma}[Solution of the Recurrence]
\label{lemma:recurrence-solution}
Let $\{a_t\}_{t=1}^\infty$ be any sequence of non-negative real numbers.
Suppose that $\{\Delta_t\}_{t=0}^\infty$ is a sequence of real numbers
satisfying
$$
\Delta_0 = 0 \qquad \text{and} \qquad
\Delta_t \le \Delta_{t-1} + \min \left\{ a_t, \ \frac{a_t^2}{\Delta_{t-1}} \right\} \quad \text{for any $t \ge 1$} \; .
$$
Then, for any $T \ge 0$,
$$
\Delta_T \le \sqrt{3 \sum_{t=1}^T a_t^2} \; .
$$
\end{lemma}
%
\begin{proof}
Observe that
\[
\Delta_T^2
=
\sum_{t=1}^T (\Delta_t - \Delta_{t-1})(\Delta_t - \Delta_{t-1} + 2\Delta_{t-1}) \; .
\]
The left term of the minimum inequality in the definition of $\Delta_t$ gives
\[
(\Delta_t - \Delta_{t-1})^2 \le a_t^2,
\]
while the right term gives
\[
2 (\Delta_t - \Delta_{t-1}) \Delta_{t-1} \le 2 a_t^2 \; .
\]
So, we conclude
\[
\Delta_T^2
\le
3 \sum_{t=1}^T a_t^2\; .
\]
\end{proof}

>>>>

In this section, we show that it is possible to derive a scale-free algorithm
different from \textsc{SOLO FTRL}. We generalize the \textsc{AdaHedge}
algorithm~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to the OLO setting,
showing that it retains its scale-free property.  We call the resulting
algorithm \textsc{AdaFTRL}. The analysis is very general and based on general
properties of strongly convex functions, rather than specific properties of the
entropic regularizer as in the original analysis of \textsc{AdaHedge}.

@@

Our main result is an $O(\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2})$ upper bound on the
regret of the algorithm after $T$ rounds, without the need to know beforehand an
upper bound on $\|\ell_t\|_*$.  We prove the theorem
in~\ref{section:ada-ftrl-regret-bound}.

@@

\begin{lemma}[Solution of the Recurrence]
\label{lemma:recurrence-solution}
Let $\{a_t\}_{t=1}^\infty$ be any sequence of non-negative real numbers.
Suppose that $\{\Delta_t\}_{t=0}^\infty$ is a sequence of non-negative
real numbers satisfying
$$
\Delta_0 = 0 \qquad \text{and} \qquad
\Delta_t \le \Delta_{t-1} + \min \left\{ a_t, \ \frac{a_t^2}{\Delta_{t-1}} \right\} \quad \text{for any $t \ge 1$} \; .
$$
Then, for any $T \ge 0$,
$$
\Delta_T \le \sqrt{3 \sum_{t=1}^T a_t^2} \; .
$$
\end{lemma}
%
\begin{proof}
Observe that
\[
\Delta_T^2
= \sum_{t=1}^T \Delta_t^2 - \Delta_{t-1}^2
= \sum_{t=1}^T (\Delta_t - \Delta_{t-1})^2 + 2 (\Delta_t - \Delta_{t-1}) \Delta_{t-1} \; .
\]
We now bound each term in the sum.
The left term of the minimum inequality in the definition of $\Delta_t$ gives
\[
(\Delta_t - \Delta_{t-1})^2 \le a_t^2,
\]
while the right term gives
\[
2 (\Delta_t - \Delta_{t-1}) \Delta_{t-1} \le 2 a_t^2 \; .
\]
So, we conclude
\[
\Delta_T^2
\le
3 \sum_{t=1}^T a_t^2\; .
\]
\end{proof}

### 2016-09-27 revision/introduction.tex
Improvements

OLO is a fundamental problem in machine
learning~\cite{Cesa-Bianchi-Lugosi-2006, Rakhlin-Sridharan-2009,
Shalev-Shwartz-2011}.  Many learning problems can be directly phrased as OLO,
e.g., learning with expert advice~\cite{Littlestone-Warmuth-1994, Vovk-1998,
Freund-Schapire-1997, Cesa-Bianchi-Haussler-Helmbold-Schapire-Warmuth-1997},
online combinatorial optimization~\cite{Kalai-Vempala-2005,
Helmbold-Warmuth-2009, Koolen-Warmuth-Kivinen-2010}. Other problems can be
reduced to OLO, e.g., online convex
optimization~\cite{Abernethy-Bartlett-Rakhlin-Tewari-2008},
\cite[Chapter~2]{Shalev-Shwartz-2011}, online classification
~\cite{Rosenblatt-1958, Freund-Schapire-1999} and
regression~\cite{Kivinen-Warmuth-1997},
~\cite[Chapters~11~and~12]{Cesa-Bianchi-Lugosi-2006}, multi-armed bandits
problems~\cite[Chapter~6]{Cesa-Bianchi-Lugosi-2006},
\cite{Abernethy-Hazan-Rakhlin-2008, Bubeck-Cesa-Bianchi-2012}, and batch and
stochastic optimization of convex functions~\cite{Nemirovski-Yudin-1983,
Bubeck-2015}.  Hence, a result in OLO immediately implies other results in all
these domains.

@@

For this reason, most of the existing OLO algorithms receive as an input---or
implicitly assume---an upper bound $B$ on the norm of the loss vectors.  The
input $B$ is often disguised as the learning rate, the regularization
parameter, or the parameter of strong convexity of the regularizer.
%Examples of
%such algorithms include \textsc{Hedge}~\cite{Freund-Schapire-1997} or online
%projected gradient descent with fixed learning rate.
However, these algorithms have two obvious drawbacks.

@@

The majority of the existing algorithms for OLO are based on two generic algorithms:
\textsc{Follow The Regularizer Leader} (\textsc{FTRL}) and \textsc{Mirror
Descent} (\textsc{MD}). \textsc{FTRL} dates back to the potential-based
forecaster in \cite[Chapter~11]{Cesa-Bianchi-Lugosi-2006} and its theory was
developed in \cite{Shalev-Shwartz-2007}.  The name \textsc{Follow The
Regularized Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}.
Independently, the same algorithm was proposed in~\cite{Nestorov-2009} for
convex optimization under the name \textsc{Dual Averaging} and rediscovered
in~\cite{Xiao-2010} for online convex optimization. Time-varying regularizers were analyzed
in~\cite{Duchi-Hazan-Singer-2011} and generalized in \cite{Orabona-Crammer-Cesa-Bianchi-2014}. \textsc{MD} was originally proposed
in \cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003}
for convex optimization. In the online learning literature it makes its first
appearance, with a different name, in~\cite{Kivinen-Warmuth-1997}.
%A good description of \textsc{MD} is given in~\cite{Rakhlin-Sridharan-2009}.

@@

Scale-free versions of \textsc{MD} include \textsc{AdaGrad
MD}~\cite{Duchi-Hazan-Singer-2011}. However, the AdaGrad
MD algorithm has a non-trivial regret bounds only when the Bregman divergence associated with the regularizer
is bounded. In particular, since a bound on the Bregman divergence implies that
the decision set is bounded, the regret bound for \textsc{AdaGrad MD} and
is vacuous for unbounded sets. In fact, as we show in
Section~\ref{subsection:mirror-descent-lower-bound}, \textsc{AdaGrad MD} and
similar algorithms based on \textsc{MD} might incur $\Omega(T)$ regret if the Bregman
divergence is not bounded.

Only one scale-free algorithm based on \textsc{FTRL} was known. It is the
\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} algorithm for
learning with expert advice, where the decision set is bounded.
An algorithm based on \textsc{FTRL} that is ``almost'' scale-free is 
\textsc{AdaGrad FTRL}~\cite{Duchi-Hazan-Singer-2011}.  This algorithm fail to
be scale-free due to ``off-by-one'' issues; see~\cite{McMahan-2014} and the
discussion in Section~\ref{section:solo-ftrl}. Instead, \textsc{FTRL-Proximal}~\cite{McMahan-Streeter-2010,McMahan-2014} solves the off-by-one issues, but it requires proximal regularizers, that in general do not have a simple form and even in the simple 2-norm case requires bounded domains.

@@

We design and analyze two scale-free algorithms: \textsc{SOLO FTRL} and \textsc{Scale-Free MD}.  A third one, \textsc{AdaFTRL}, is present in the Appendix~\ref{section:ada-ftrl}. \textsc{AdaFTRL} and
\textsc{SOLO FTRL} are based on \textsc{FTRL}.  \textsc{AdaFTRL} is a
generalization of
\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to arbitrary
strongly convex regularizers.  \textsc{SOLO FTRL} can be viewed as the
``correct'' scale-free version of the diagonal version of \textsc{AdaGrad
FTRL}~\cite{Duchi-Hazan-Singer-2011} generalized to arbitrary strongly convex
regularizers.  \textsc{Scale-Free MD} is based on \textsc{MD}. It is a
generalization of \textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011} to
arbitrary strongly convex regularizers.  The three algorithms are presented in
Sections~\ref{section:ada-ftrl}, \ref{section:solo-ftrl} and
\ref{section:mirror-descent} respectively.

We prove that the regret of \textsc{SOLO FTRL} and \textsc{AdaFTRL} after $T$
rounds is bounded by $O (\sqrt{\sup_{v \in K} f(v)
\sum_{t=1}^T\|\ell_t\|_*^2} )$ where $f$ is a non-negative regularizer
that is $1$-strongly convex with respect to a norm $\|\cdot\|$ and
$\|\cdot\|_*$ is its dual norm. For \textsc{Scale-Free MD}, we prove $O
(\sqrt{\sup_{u,v \in K} B_f(u,v) \sum_{t=1}^T\|\ell_t\|_*^2} )$
where $B_f$ is the Bregman divergence associated with a $1$-strongly convex
regularizer $f$.  
%However, as we discuss in Section~\ref{section:mirror-descent}, the
%upper bound for \textsc{Scale-Free MD} is never better than the bound
%for \textsc{SOLO FTRL} and \textsc{AdaFTRL}.
In Section~\ref{section:lower-bound}, we show that the $\sqrt{\sum_{t=1}^T
\|\ell_t\|_*^2}$ term in the bounds is necessary by proving a $\frac{D}{\sqrt{8}}
\sqrt{\sum_{t=1}^T\|\ell_t\|_*^2}$ lower bound on the regret of any algorithm
for OLO for any decision set with diameter $D$ with respect to the primal norm
$\|\cdot\|$.

@@

Finally, in Section~\ref{subsection:mirror-descent-lower-bound}, we show
negative results for existing popular variants of \textsc{MD}. We show two
examples that demonstrate that variants of \textsc{MD} can have $\Omega(T)$
regret (on sequences of loss vectors of unit norm) if the Bregman divergence
associated with the regularizer is unbounded.
%First example demonstrates
%$\Omega(T^{3/2})$ regret of the standard gradient descent with step size
%$1/\sqrt{t}$, \textsc{AdaGrad MD} and \textsc{Scale-Free MD} on a unbounded
%decision set.  Second example demonstrates $\Omega(T)$ regret for
%\textsc{Scale-Free MD} for learning with expert advice with negative entropy
%regularizer.
These results indicate that \textsc{FTRL} is superior to \textsc{MD}.

>>>>

OLO is a fundamental problem in machine
learning~\cite{Cesa-Bianchi-Lugosi-2006, Rakhlin-Sridharan-2009,
Shalev-Shwartz-2011}.  Many learning problems can be directly phrased as OLO,
e.g., learning with expert advice~\cite{Littlestone-Warmuth-1994, Vovk-1998,
Freund-Schapire-1997, Cesa-Bianchi-Haussler-Helmbold-Schapire-Warmuth-1997}
and online combinatorial optimization~\cite{Kalai-Vempala-2005,
Helmbold-Warmuth-2009, Koolen-Warmuth-Kivinen-2010}. Other problems can be
reduced to OLO, e.g., online convex
optimization~\cite{Abernethy-Bartlett-Rakhlin-Tewari-2008},
\cite[Chapter~2]{Shalev-Shwartz-2011}, online classification
~\cite{Rosenblatt-1958, Freund-Schapire-1999} and
regression~\cite{Kivinen-Warmuth-1997},
~\cite[Chapters~11~and~12]{Cesa-Bianchi-Lugosi-2006}, multi-armed bandits
problems~\cite[Chapter~6]{Cesa-Bianchi-Lugosi-2006},
\cite{Abernethy-Hazan-Rakhlin-2008, Bubeck-Cesa-Bianchi-2012}, and batch and
stochastic optimization of convex functions~\cite{Nemirovski-Yudin-1983,
Bubeck-2015}.  Hence, a result in OLO immediately implies other results in all
these domains.

@@

For this reason, most of the existing OLO algorithms receive as an input---or
implicitly assume---an upper bound $B$ on the norm of the loss vectors.  The
input $B$ is often disguised as the learning rate, the regularization
parameter, or the parameter of strong convexity of the regularizer.
However, these algorithms have two obvious drawbacks.

@@

The majority of the existing algorithms for OLO are based on two generic
algorithms: \textsc{Follow The Regularizer Leader} (\textsc{FTRL}) and
\textsc{Mirror Descent} (\textsc{MD}). \textsc{FTRL} dates back to the
potential-based forecaster in \cite[Chapter~11]{Cesa-Bianchi-Lugosi-2006} and
its theory was developed in \cite{Shalev-Shwartz-2007}.  The name \textsc{Follow
The Regularized Leader} comes from~\cite{Abernethy-Hazan-Rakhlin-2008}.
Independently, the same algorithm was proposed in~\cite{Nestorov-2009} for
convex optimization under the name \textsc{Dual Averaging} and rediscovered
in~\cite{Xiao-2010} for online convex optimization. Time-varying regularizers
were analyzed in~\cite{Duchi-Hazan-Singer-2011} and generalized in
\cite{Orabona-Crammer-Cesa-Bianchi-2014}. \textsc{MD} was originally proposed in
\cite{Nemirovski-Yudin-1983} and later analyzed in~\cite{Beck-Teboulle-2003} for
convex optimization. In the online learning literature it makes its first
appearance, with a different name, in~\cite{Kivinen-Warmuth-1997}.

@@

Scale-free versions of \textsc{MD} include \textsc{AdaGrad
MD}~\cite{Duchi-Hazan-Singer-2011}. However, the AdaGrad MD algorithm has a
non-trivial regret bounds only when the Bregman divergence associated with the
regularizer is bounded. In particular, since a bound on the Bregman divergence
implies that the decision set is bounded, the regret bound for \textsc{AdaGrad
MD} and is vacuous for unbounded sets. In fact, as we show in
Section~\ref{subsection:mirror-descent-lower-bound}, \textsc{AdaGrad MD} and
similar algorithms based on \textsc{MD} incurs $\Omega(T)$ regret, in the worst
case, if the Bregman divergence is not bounded.

Only one scale-free algorithm based on \textsc{FTRL} was known. It is the
\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} algorithm for
learning with expert advice, where the decision set is bounded. An algorithm
based on \textsc{FTRL} that is ``almost'' scale-free is \textsc{AdaGrad
FTRL}~\cite{Duchi-Hazan-Singer-2011}.  This algorithm fail to be scale-free due
to ``off-by-one'' issue; see~\cite{McMahan-2014} and the discussion in
Section~\ref{section:solo-ftrl}. Instead,
\textsc{FTRL-Proximal}~\cite{McMahan-Streeter-2010,McMahan-2014} solves the
off-by-one issue, but it requires proximal regularizers, that in general do not
have a simple form and even in the simple 2-norm case requires bounded domains.

@@

We design and analyze two scale-free algorithms: \textsc{SOLO FTRL} and
\textsc{Scale-Free MD}.  A third one, \textsc{AdaFTRL}, is presented in
the Appendix. \textsc{SOLO FTRL} and \textsc{AdaFTRL} are based on
\textsc{FTRL}.  \textsc{AdaFTRL} is a generalization of
\textsc{AdaHedge}~\cite{de-Rooij-van-Erven-Grunwald-Koolen-2014} to arbitrary
strongly convex regularizers.  \textsc{SOLO FTRL} can be viewed as the
``correct'' scale-free version of the diagonal version of \textsc{AdaGrad
FTRL}~\cite{Duchi-Hazan-Singer-2011} generalized to arbitrary strongly convex
regularizers.  \textsc{Scale-Free MD} is based on \textsc{MD}. It is a
generalization of \textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011} to
arbitrary strongly convex regularizers.  The three algorithms are presented in
Sections~\ref{section:solo-ftrl} and \ref{section:mirror-descent}, and
\ref{section:ada-ftrl}, respectively.

We prove that the regret of \textsc{SOLO FTRL} and \textsc{AdaFTRL} after $T$
rounds is bounded by $O (\sqrt{\sup_{v \in K} f(v) \sum_{t=1}^T\|\ell_t\|_*^2} )$
where $f$ is a non-negative regularizer that is $1$-strongly convex with respect
to a norm $\|\cdot\|$ and $\|\cdot\|_*$ is its dual norm. For \textsc{Scale-Free
MD}, we prove $O (\sqrt{\sup_{u,v \in K} B_f(u,v) \sum_{t=1}^T\|\ell_t\|_*^2} )$
where $B_f$ is the Bregman divergence associated with a $1$-strongly convex
regularizer $f$. In Section~\ref{section:lower-bound}, we show that the
$\sqrt{\sum_{t=1}^T \|\ell_t\|_*^2}$ term in the bounds is necessary by proving
a $\frac{D}{\sqrt{8}} \sqrt{\sum_{t=1}^T\|\ell_t\|_*^2}$ lower bound on the
regret of any algorithm for OLO for any decision set with diameter $D$ with
respect to the primal norm $\|\cdot\|$.

@@

Finally, in Section~\ref{subsection:mirror-descent-lower-bound}, we show
negative results for existing popular variants of \textsc{MD}. We show two
examples that demonstrate that variants of \textsc{MD} can have $\Omega(T)$
regret (on sequences of loss vectors of unit norm) if the Bregman divergence
associated with the regularizer is unbounded. These results indicate that
\textsc{FTRL} is superior to \textsc{MD}.

### 2016-09-27 revision/mirror-descent.tex
Improvements

Full-matrix and diagonal versions of \textsc{AdaGrad with Composite MD
Update}~\cite{Duchi-Hazan-Singer-2011} are instances of \textsc{MD} for any
bounded $K$ with regularizers
$$
R_t(w) = \frac{1}{2} w^\top \left(\sum_{i=1}^t \ell_i \ell_i^\top \right)^{1/2} \!\!\!\! w
\qquad \text{and} \qquad
R_t(w) = \frac{1}{2} \sum_{j=1}^d w_j^2 \sqrt{ \sum_{i=1}^t \ell_{i,j}^2}
$$
respectively. The diagonal version is simply the per-coordinate construction
applied to the one-dimensional version, see Section~\ref{sec:per-coordinate}. When $d=1$, both algorithms are
equivalent to
\begin{equation}
\label{equation:ada-grad-regularizer}
R_t(w) = \frac{\|w\|_2^2}{2} \sqrt{\sum_{i=1}^t \|\ell_i\|_2^2 } \; .
\end{equation}
All three sequences of regularizers give a scale-free algorithm. Regularizer
\eqref{equation:ada-grad-regularizer} can be viewed as a scale-free
generalization of GIGA. That is, when $\|\ell_i\|_2
= 1$ for all $i$, \textsc{MD} with regularizer
\eqref{equation:ada-grad-regularizer} coincides with GIGA.

Replacing $\frac{1}{2}\|w\|_2^2$ in equation
\eqref{equation:ada-grad-regularizer} with an arbitrary strongly convex
function $R:K \to \R$ yields
\begin{equation}
\label{equation:scale-free-mirror-descent}
R_t(w) = R(w) \sqrt{\sum_{i=1}^t \|\ell_i\|_*^2} \; .
\end{equation}
We call the resulting algorithm \textsc{Scale-Free MD} with
regularizer $R$. Similar to \textsc{SOLO FTRL}, the regularizer
\eqref{equation:scale-free-mirror-descent} does not uniquely define
the \textsc{MD} minimizer
$w_{t+1} = \argmin_{w \in K} \left(\langle \ell_t, w \rangle + \Breg_{R_t}(w,
w_t) \right)$ when $\sqrt{\sum_{i=1}^t \|\ell_i\|_*^2}$ is zero.
This happens when the loss vectors $\ell_1, \ell_2, \dots, \ell_t$ are all
zero. In this case, we define
$w_{t+1} = \argmin_{w \in K} R(w)$ which agrees with
$w_{t+1} = \lim_{a \to 0^+} \argmin_{w \in K} a \Breg_{R}(w, w_t)$.
Similarly, $w_1 = \argmin_{w \in K} R(w)$.

>>>>

In this section, we analyze scale-free version of \textsc{Mirror Descent}.
Our algorithm uses regularizer
\begin{equation}
\label{equation:scale-free-mirror-descent}
R_t(w) = R(w) \sqrt{\sum_{i=1}^t \|\ell_i\|_*^2} \; .
\end{equation}
where $R:K \to \R$ an arbitrary strongly convex function.
We call the resulting algorithm \textsc{Scale-Free MD}.
Similar to \textsc{SOLO FTRL}, the regularizer
\eqref{equation:scale-free-mirror-descent} does not uniquely define
the \textsc{MD} minimizer
$w_{t+1} = \argmin_{w \in K} \left(\langle \ell_t, w \rangle + \Breg_{R_t}(w,
w_t) \right)$ when $\sqrt{\sum_{i=1}^t \|\ell_i\|_*^2}$ is zero.
This happens when the loss vectors $\ell_1, \ell_2, \dots, \ell_t$ are all
zero. In this case, we define
$w_{t+1} = \argmin_{w \in K} R(w)$ which agrees with
$w_{t+1} = \lim_{a \to 0^+} \argmin_{w \in K} a \Breg_{R}(w, w_t)$.
Similarly, $w_1 = \argmin_{w \in K} R(w)$.

Per-coordinate version of \textsc{Scale-Free MD} with regularizer $R(w) =
\frac{1}{2}\norm{w}_2^2$ is exactly the same as the diagonal version of
\textsc{AdaGrad MD}~\cite{Duchi-Hazan-Singer-2011}.

### 2016-09-27 revision/preliminaries.tex
Improvements

\textsc{Mirror Descent} (MD) is a generic algorithm similar to \textsc{FTRL}
but quite different in the details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
not differentiable,\footnote{Note that this can happen even when $R_t$ is a
restriction of a differentiable function defined on a superset of $K$.  If $K$
is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
If $K$ is a subset of an affine subspace of a dimension smaller than the
dimension of $V$, then $R_t$ fails to be differentiable everywhere.} the
Bregman divergence, $\Breg_{R_t}(u,v) = R_t(u) - R_t(v) - \langle \grad R_t(v),
u - v \rangle$ needs to be defined. This is done by choosing a subgradient map
$\grad R_t:K \to V$, i.e., a function such that $\grad R_t(w)$ is a subgradient
of $R_t$ at any point $w$. If $R_t$ is a restriction of a differentiable function
$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
\in K$. The following lemma bounds the regret of \textsc{MD}.  A proof can be found
in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
completeness we give a proof in~\ref{section:mirror-descent-proofs}.

@@

Several existing algorithms are special cases of \textsc{MD}.  For example,
\textsc{Hedge} is an instance of \textsc{MD} for the probability simplex $K =
\{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy
regularizer $R_t(w) = \frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $\eta > 0$
is a (constant) learning rate. Note that this algorithm is \emph{not}
scale-free.

The most prominent instance of \textsc{MD} is \textsc{Projected Gradient
Descent} defined by the recurrence
$$
w_{t+1} = \Pi_K \left( w_t - \eta_t \ell_t \right) \; ,
$$
where $\Pi_K(x) = \argmin_{y \in K} \|x-y\|_2$ is the Euclidean projection to
$K$ and $\eta_t > 0$ is the step size.  \textsc{Projected Gradient Descent} is
an instance of \textsc{MD} with regularizer $R_t(w) =
\frac{1}{2\eta_t}\|w\|_2^2$.  A special case of \textsc{Projected Gradient
Descent} with step size $\eta_t = \frac{1}{\sqrt{t}}$ is \textsc{Generalized
Infinitesimal Gradient Ascent} (GIGA)~\cite{Zinkevich-2003}.

>>>>

\textsc{Mirror Descent} (MD) is a generic algorithm similar to \textsc{FTRL}
but quite different in the details. The algorithm is stated as
Algorithm~\ref{algorithm:mirror-descent-varying-regularizer}. The algorithm is
parametrized by a sequence $\{R_t\}_{t=0}^\infty$ of convex functions $R_t:K
\to \R$ called \emph{regularizers}. Each regularizer $R_t$ can depend on past
loss vectors $\ell_1, \ell_2, \dots, \ell_t$ in an arbitrary way. If $R_t$ is
not differentiable,\footnote{Note that this can happen even when $R_t$ is a
restriction of a differentiable function defined on a superset of $K$.  If $K$
is bounded and closed, $R_t$ fails to be differentiable at the boundary of $K$.
If $K$ is a subset of an affine subspace of a dimension smaller than the
dimension of $V$, then $R_t$ fails to be differentiable everywhere.} the
Bregman divergence, $\Breg_{R_t}(u,v) = R_t(u) - R_t(v) - \langle \grad R_t(v),
u - v \rangle$ needs to be defined. This is done by choosing a subgradient map
$\grad R_t:K \to V$, i.e., a function such that $\grad R_t(w)$ is a subgradient
of $R_t$ at any point $w$. If $R_t$ is a restriction of a differentiable function
$R'_t$, it is convenient to define $\grad R_t(w) = \grad R'_t(w)$ for all $w
\in K$. The following lemma bounds the regret of \textsc{MD}.

@@

The proof of the lemma can be found
in~\cite{Rakhlin-Sridharan-2009,Duchi-Shalev-Shwartz-Singer-Tewari-2010}.  For
completeness, we give a proof in~\ref{section:mirror-descent-proofs}.

%Several existing algorithms are special cases of \textsc{MD}.  For example,
%\textsc{Hedge} is an instance of \textsc{MD} for the probability simplex $K =
%\{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$ with negative entropy
%regularizer $R_t(w) = \frac{1}{\eta} \sum_{j=1}^d w_j \ln w_j$ where $\eta > 0$
%is a (constant) learning rate. Note that this algorithm is \emph{not}
%scale-free.

%The most prominent instance of \textsc{MD} is \textsc{Projected Gradient
%Descent} defined by the recurrence
%$$
%w_{t+1} = \Pi_K \left( w_t - \eta_t \ell_t \right) \; ,
%$$
%where $\Pi_K(x) = \argmin_{y \in K} \|x-y\|_2$ is the Euclidean projection to
%$K$ and $\eta_t > 0$ is the step size.  \textsc{Projected Gradient Descent} is
%an instance of \textsc{MD} with regularizer $R_t(w) =
%\frac{1}{2\eta_t}\|w\|_2^2$.  A special case of \textsc{Projected Gradient
%Descent} with step size $\eta_t = \frac{1}{\sqrt{t}}$ is \textsc{Generalized
%Infinitesimal Gradient Ascent} (GIGA)~\cite{Zinkevich-2003}.
